---
title: "BRAZILIAN COFFEE PRODUCTION (1961-2015)"
author: "Andriy Gorshkov"
date: "18 Dezember 2018"
output: 
  html_document:
    toc: TRUE
    toc_float: yes
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(ggplot2)
library(dplyr)
library(readxl)
library(knitr)
library(lmtest)
library(car)
library(data.table)

#DATA
rain <- read_xls("pr_1901_2015.xls")
temp <- read_xls("tas_1901_2015.xls")
br_crop <- as_data_frame(read.csv("FAOSTAT_data_12-19-2018.csv"))
br_crop <- br_crop %>% 
  filter((Unit == "ha" | Unit == "tonnes") & Year <= 2015)

#Fitting
rain_temp <- rain %>% 
  inner_join(temp, by = c("Year", "Month")) %>% 
  select(Year, Month, pr, t_cel=tas) %>% 
  filter(Year %in% seq(from=1961, to=max(as.integer(br_crop$Year))))
rain_temp$Year <- as.factor(rain_temp$Year)

br_crop_ha <- br_crop %>% 
  filter(Unit == "ha") 

br_crop_t <- br_crop %>% 
  filter(Unit == "tonnes") 

br_crop_cum <- br_crop_ha %>% 
  inner_join(br_crop_t, by = c("Year")) %>% 
  select(Year, ha=Value.x, tonnes=Value.y)

br_crop_cum$Year <- as.character(br_crop_cum$Year)

rain_temp_crop <- rain_temp %>% 
  group_by(Year) %>% 
  summarise(pr = mean(pr), t_cel = mean(t_cel)) %>% 
  inner_join(br_crop_cum, by = c("Year")) %>% 
  select(Year, pr, t_cel, ha, tonnes)

lmp <- function (modelobject) {
 if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
 f <- summary(modelobject)$fstatistic
 p <- pf(f[1],f[2],f[3],lower.tail=F)
 attributes(p) <- NULL
 return(p)
}

```

## Production
Let's look at Diagram 1 of the production of coffee between 1960 and 2015 crops years.
Here you can see the productivity (1000 tonnes) and harvested area (1000 ha)[^1].

[^1]: [FAO Stat](http://www.fao.org/faostat/en/#data/QC) 

```{r br_weather_m, echo=FALSE, warning=FALSE}
ggplot(br_crop, aes(x=as.factor(Year), y = Value/1000)) +
  geom_col(aes(fill = Unit), position = "dodge") +
  scale_x_discrete(breaks = as.factor(seq(from=1960, to=max(as.integer(br_crop$Year)),by = 5)))+
  geom_vline(xintercept = 42, color = "blue", size = 1, alpha = 0.9, linetype = "longdash") +
  geom_vline(xintercept = 52, color = "blue", size = 1, alpha = 0.9, linetype = "longdash") +
  xlab("Year") +
  ylab("Thousand units") +
  ggtitle("Diagram 1, Area Harvested (1000 ha) / Production (1000 t)")+
  theme_classic()
```

## Weather
Of course, the significant effect for the coffee crop makes the weather (Diagram 2 and Diagram 3)[^2]:

[^2]: [Climate Change Knowledge Portal](http://sdwebx.worldbank.org/climateportal/index.cfm?page=downscaled_data_download&menu=historical) 

```{r pressure, echo=FALSE, warning=FALSE}
ggplot(rain_temp) +
  geom_boxplot( aes(x=Year, y=pr), fill = "blue", alpha = 0.5) +
  scale_x_discrete(breaks = as.factor(seq(from=1960, to=max(as.integer(br_crop$Year)),by = 5))) +
  ylab("mm") +
  ggtitle("Diagram 2, Precipitation(mm)")+
  theme_classic()

ggplot(rain_temp) +
  geom_boxplot( aes(x=Year, y=t_cel), fill = "red", alpha = 0.5) +
  scale_x_discrete(breaks = as.factor(seq(from=1960, to=max(as.integer(br_crop$Year)),by = 5))) +
  ggtitle("Diagram 3, Temperature(t° C)")+
  ylab("t° C") +
  theme_classic()
```

## Linear models

The dependent variable in our model is Production in tonns.
Predictor-variables: harvested area (ha), Precipitation (pr) and temperature (t_cel).

###Normality test
Quantile-Comparison Plots and Shapiro-Wilk normality tests:
```{r Normality1, echo=FALSE, message=FALSE, warning=FALSE}
qq <- qqPlot(rain_temp_crop$ha, main = "Harvested area (ha)", ylab = "",  xlab = "")
```

p-value = `r shapiro.test(rain_temp_crop$ha)$p.value` If the P-value less then 0.05 then it means that we have got not a normal distribution for observation. But another side, if we exclude some of the observations, Quantile-Comparison Plot show us enough normal distribution.

```{r Normality2, echo=FALSE, message=FALSE, warning=FALSE}
qq <- qqPlot(rain_temp_crop$pr, main = "Precipitation (pr)", ylab = "",  xlab = "")
```

p-value = `r shapiro.test(rain_temp_crop$pr)$p.value`

```{r Normality3, echo=FALSE, message=FALSE, warning=FALSE}
qq <- qqPlot(rain_temp_crop$t_cel, main = "Temperature (t_cel)", ylab = "",  xlab = "")
```

p-value = `r shapiro.test(rain_temp_crop$t_cel)$p.value`

###Lack of perfect multicollinearity in the predictors test:

```{r multicollinearity, echo=FALSE, warning=FALSE}
pairs(rain_temp_crop[c(-1,-5)])
```

Obviously, there is no multicollinearity in the predictors.

###Durbin-Watson test for autocorrelation of disturbances:
```{r dwtest, echo=FALSE, message=TRUE, warning=FALSE}
dwtest(tonnes ~ pr+t_cel+ha, data = rain_temp_crop)
```

###Breusch-Pagan test against heteroskedasticity:
```{r bptest, echo=FALSE, message=TRUE, warning=FALSE}
bptest(tonnes ~ pr+t_cel+ha, data = rain_temp_crop)
```


###Model 1
Let's try to fit the model without interaction between the predictors (Table 1,  Model 1):
```{r LinearModel_1_1, echo=FALSE, warning=FALSE}
r_model_1 <- lm(data=rain_temp_crop,  tonnes ~ pr+t_cel+ha)
s_1 <- summary(r_model_1)
a_1 <- anova(r_model_1)
kable(s_1$coefficients, caption = paste("Table 1, Model 1", Reduce(paste, deparse(s_1$terms))))
```

Adjusted R-squared: `r s_1$adj.r.squared`

P-Value: `r lmp(r_model_1)`

###Model 2
and with an interaction between the Precipitation (pr) and the temperature (t_cel) predictors (Table 2,  Model 2):
```{r LinearModel_1_2, echo=FALSE, warning=FALSE}
r_model_2 <- lm(data=rain_temp_crop,  tonnes ~ pr*t_cel+ha)
s_2 <- summary(r_model_2)
a_2 <- anova(r_model_2)
kable(s_2$coefficients, caption = paste("Table 2,  Model 2", Reduce(paste, deparse(s_2$terms))))
```

Adjusted R-squared: `r s_2$adj.r.squared`

P-Value: `r lmp(r_model_2)`

###Prediction overview
The adjusted R-squared and the P-Values of both models are very close to each other. So maybe it makes no sense to complicate the model.
Let's build predictions based on these two models and show them on Diagram 4: 

```{r Predictions, echo=FALSE, warning=FALSE}
rain_temp_crop$p_tonn_model1 <- predict(r_model_1)
rain_temp_crop$p_tonn_model2 <- predict(r_model_2)

ggplot(rain_temp_crop, aes(as.integer(Year)))+
  geom_line(aes(y = tonnes/1000, color = "Real value, tons")) +
  geom_line(aes(y = p_tonn_model1/1000, color = "Model 1")) +
  geom_line(aes(y = p_tonn_model2/1000, color = "Model 2")) +
  geom_vline(xintercept = 2011, color = "blue", size = 0.5, alpha = 0.9, linetype = "longdash") +
  #geom_vline(xintercept = 2015, color = "blue", size = 0.5, alpha = 0.9, linetype = "longdash") +
  scale_x_continuous(breaks = (seq(from=1960, to=max(as.integer(br_crop$Year)),by = 5)))+
  ggtitle("Diagram 4, Comparison of predictions") +
  xlab("Year") +
  ylab("Thousand tons") +  
  theme_classic()
```

We can note that substantial productivity was significantly increased between 2002 and 2012 (Diagram 1) and our predictions works after that not so good as before. But I assume, that after the boom of the technical progress the models will show better result or we need to include another predictor, like Investing in the agriculture sector. 

##Improving model
Maybe we can improve the model if we define the most important months for crop. But let's analyze, how to impact previous year for the crop of the current year.

```{r DataForImproving, echo=FALSE, warning=FALSE}
temp_by_month <- as_data_frame(dcast(temp , Year~Month, value.var = "tas"))
rain_by_month <- as_data_frame(dcast(rain , Year~Month, value.var = "pr"))

rain_by_month$Year <- as.character(as.integer(rain_by_month$Year) - 1)
temp_by_month$Year <- as.character(as.integer(temp_by_month$Year) - 1)

kable(head(temp_by_month,5),caption = "temperature")
kable(head(rain_by_month,5),caption = "Precipitation")
#rain_by_month$Year <- as.character(as.integer(rain_by_month$Year) - 1)

twelve_m_t <-  br_crop_cum %>% 
  inner_join(temp_by_month, c("Year")) 

twelve_m_p <-  br_crop_cum %>% 
  inner_join(rain_by_month, c("Year"))
#"Model for temperature, tonnes ~ ."
r_model <- lm(data=twelve_m_t[-1],  tonnes ~ .)
summary(r_model)
#"Model for Precipitation, tonnes ~ ."
r_model <- lm(data=twelve_m_p[-1],  tonnes ~ .)
summary(r_model)
```

So, let's use the month 6 with observation for "temperature" and 2, 5, 9, 10 with observation for precipitation.
And we got follow table for the model:

```{r DataForImproving1, echo=FALSE, warning=FALSE}
twelve_m <-  br_crop_cum %>% 
  inner_join(temp_by_month, c("Year")) %>% 
  inner_join(rain_by_month, c("Year")) %>% 
  select(Year, ha, tonnes, t6 = `6.x`, p2p = `2.y`, p5p=`5.y`, p9p=`9.y`, p10p=`10.y`)
  #select(Year, ha, tonnes, t1 = `1.x`, t4 = `4.x`, t7 = `7.x`, t9 = `9.x`, r2p=`9.y`)

kable(head(twelve_m))
```

###Improved model

```{r DataForImproving3, echo=FALSE, warning=FALSE}
r_model3 <- lm(data=twelve_m[-1],  tonnes ~ .)
summary(r_model3)

rain_temp_crop <- rain_temp_crop %>% 
  left_join(twelve_m, c("Year")) %>% 
  select(Year, pr, t_cel, ha = ha.x, tonnes = tonnes.x, t6,p2p, p5p, p9p, p10p, p_tonn_model1, p_tonn_model2)

rain_temp_crop$p_tonn_model3 <- predict(r_model3, rain_temp_crop)

ggplot(rain_temp_crop, aes(as.integer(Year)))+
  geom_line(aes(y = tonnes/1000, color = "Real value, tons"), size = 1) +
  geom_line(aes(y = p_tonn_model1/1000, color = "Model 1")) +
  #geom_line(aes(y = p_tonn_model2/1000, color = "Model 2")) +
  geom_line(aes(y = p_tonn_model3/1000, color = "Improved model"), size = 1) +
  geom_vline(xintercept = 2011, color = "blue", size = 0.5, alpha = 0.9, linetype = "longdash") +
  #geom_vline(xintercept = 2015, color = "blue", size = 0.5, alpha = 0.9, linetype = "longdash") +
  scale_x_continuous(breaks = (seq(from=1960, to=max(as.integer(br_crop$Year)),by = 5)))+
  ggtitle("Diagram 5, Comparison of predictions") +
  xlab("Year") +
  ylab("Thousand tons") +  
  theme_classic()
```


